---
title: "Data-Pipeline"
author: "Claire McLean"
date: "2024-10-28"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
NOTE: The dataset I am using for this project is named basque_parliament_1 and was found on https://huggingface.co. The original project was funded by the Spanish Ministry of Science and Innovation, and is licensed as cc0-1.0.
Citation:
@misc {software_technologies_working_group_2024,
author       = { {Software Technologies Working Group} },
title        = { basque_parliament_1 (Revision a2fbcaf) },
year         = 2024,
url          = { https://huggingface.co/datasets/gttsehu/basque_parliament_1 },
doi          = { 10.57967/hf/2485 },
publisher    = { Hugging Face }
}

# Data Pipeline

## Step 1: Importing major dependencies

First, import dependencies:
```{r}
library("tidyverse")
library("readr")
library("tidytext")
library("dplyr")
```

## Step 2: The following code chunks are the pipeline I followed to split the data into two smaller dataframes. This pipeline can be ignored.

Read in the .tsv file:
```{r}
#raw_data <- read_tsv("train.tsv")
#raw_data
```

Finding number of total rows in the raw data:
```{r}
#row_count <- nrow(raw_data)
#row_count
```

Now, breaking up the dataset into two separate parts so that it is possible to push to GitHub (100MB maximum):
```{r}
#row_split1 <- (nrow(raw_data)/2) %>% round(digits = 0)
#row_split2 <- row_split1+1
#raw_data1 <- raw_data %>%
  #slice(1:row_split)
#raw_data2 <- raw_data %>%
  #slice(row_split2:row_count)
```

Exporting the two above datasets into .csv files:
```{r}
#write_csv(raw_data1, "raw_data1.csv")
#write_csv(raw_data2, "raw_data2.csv")
```


## Step 3: Uploading raw .csv files:
```{r}
raw_data1 <- read_csv("raw_data1.csv")
raw_data2 <- read_csv("raw_data2.csv")
```

Combining the two dataframes using rbind():
```{r}
raw_data <- rbind(raw_data1, raw_data2)
```

Viewing new, combined dataframe:
```{r}
nrow(raw_data)
ncol(raw_data)
```
Starting with 749,945 rows/utterances

## Step 4: Data Cleaning

Now, filtering to only include rows that are Spanish utterances (exclude Basque and bilingual):
```{r}
small_data_span <- raw_data %>% filter(language == "es")
nrow(small_data_span)
```
Now 524,942 rows/utterances (lost 225,021)

Using stringr to identify utterances/sentences that contain the following keywords. ['hablante', 'euskera', 'lenguaje', 'idioma', 'lingüístic']. This will allow me to perform sentiment analysis only on the utterances that are related in some way to language. Specifically, using str_detect() will allow me to identify occurrences of both 'hablante/es', 'lenguaje/es', 'lingüística/o/as/os'. Further, I created another vector that is used to detect occurrences of ['castellan', 'galleg'], which detects both 'castellano/a' and 'gallego/a'.
```{r}
keywords_vec <- ('hablante|euskera|lenguaje|idioma|lingüístic')
keywords_vec1 <- ('castellan|galleg')
keywords_col <- small_data_span$sentence %>%
  str_detect(keywords_vec)

keywords_col2 <- small_data_span$sentence %>% 
  str_detect(keywords_vec1)

small_data_span2 <- small_data_span %>%
  mutate(keywords = keywords_col) %>% 
  mutate(non_keywords = keywords_col2)

small_data_span3 <- small_data_span2 %>%
  filter(keywords == TRUE) %>% 
  filter(non_keywords == FALSE)

nrow(small_data_span3)
```
Now 3209 rows (lost 521,715 rows)

Changing name of dataframe:
```{r}
sampled_data <- small_data_span3
```

## Some basic stats of cleaned Parliament data:
```{r}
#number of rows:
nrow(sampled_data)
#number of columns:
ncol(sampled_data)
#average utterance length
mean(sampled_data$length)
#number of unique speakers
n_distinct(sampled_data$speaker_id)
```

## Step 5: Wrangling SEL lexicon

Load dependencies:
```{r}
library("readxl")
library("readr")
```

Reading in Spanish Emotion Lexicon file:
```{r}
sel <- read_excel("SEL.xlsx")
nrow(sel)
```
Starting with 2,036 emotion words in the SEL.

Viewed final plots and assessed which words carried less emotional 'weight'. Removed these words from the dataframe as they do not contribute much to the analysis. Additionally, created word stemmer to isolate emotion words without gender suffix, as all gendered words in the dataset are masculine:
```{r}
library("SnowballC")

low_ew_vec <- ('cuent|solo|sola|llev|lleg|sab|dec|dich|ped|mano')
low_ew <- sel$Palabra %>% 
  str_detect(low_ew_vec)
low_ew2 <- sel %>% 
  mutate('LowEmotion' = low_ew)
sel2 <- sel %>% 
  filter(low_ew == FALSE)

stemmed_sel <- wordStem(sel2$Palabra, language = "spanish")
stemmed_sel <- sel2 %>% 
  mutate(Stems = stemmed_sel) %>% select('#', Stems, PFA, Categoría, -Palabra)
nrow(stemmed_sel)

#If stem occurs multiple times in same emotion, delete duplicates. If stem occurs multiple times across different emotions, keep duplicates:
stemmed_sel <- stemmed_sel %>% group_by(Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)
nrow(stemmed_sel)
```
Now, 1,548 tokens in SEL (deleted all occurrences of words with low emotional weight, and all duplicates of stems that occurred within the same emotion)

Finding number of emotion words per emotion:
```{r}
happiness <- stemmed_sel %>% 
  filter(Categoría=='Alegría') %>% 
  nrow()
anger <- stemmed_sel %>% 
  filter(Categoría=='Enojo') %>% 
  nrow()
fear <- stemmed_sel %>% 
  filter(Categoría=='Miedo') %>% 
  nrow()
disgust <- stemmed_sel %>% 
  filter(Categoría=='Repulsión') %>% 
  nrow()
surprise <- stemmed_sel %>% 
  filter(Categoría=='Sorpresa') %>% 
  nrow()
sadness <- stemmed_sel %>% 
  filter(Categoría=='Tristeza') %>% 
  nrow()

sel_distribution <- stemmed_sel %>%
  group_by(Categoría) %>% 
  mutate(Freq=n()) %>%
  select(Categoría, Freq) %>% 
  distinct()
sel_distribution
```
## Step 6: Tokenizing and stemming Parliament sentence data

Tokenizing sentences (ngrams):
```{r}
library(dplyr)
library(tidytext)

sampled_data_ngrams <- sampled_data %>%
  unnest_tokens(tokens, sentence, token="words", drop=FALSE, to_lower=TRUE)
```

Stemming sentence tokens (ngrams):
```{r}
token_stems <- wordStem(sampled_data_ngrams$tokens, language = "spanish")
stemmed_ngrams <- sampled_data_ngrams %>% 
  mutate(Stems = token_stems)
nrow(stemmed_ngrams)
```
66,078 stems total (before inner-joining with SEL)

## Step 7: Merging Parliament data and SEL data

The following two code chunks were inspired by a project by Kayleah Griffen, which can be found at the following site: https://rpubs.com/klgriffen96/data607_hw10

Does word correspond to a particular emotion?
```{r}
merged_data <- stemmed_ngrams %>% 
  inner_join(stemmed_sel, by="Stems", multiple = "all", unmatched = "drop", relationship = "many-to-many")

nrow(merged_data)
```
2422 stems after inner-joining with SEL (stems that did not have matches in the SEL were dropped)


Grouping by sentence/utterance:
```{r}
merged_data2 <- merged_data %>% 
  arrange(speaker_id, group_by=TRUE)
```

Creating wordclouds for each sentiment:
```{r}
library(wordcloud)
library(RColorBrewer)

#Ungrouping to create wordclouds
merged_data3 <- merged_data %>% 
  ungroup()
```

Below, I am grouping each emotion and its corresponding tokens into an object. In addition to grouping them, I also deleted duplicate stems that would otherwise present problems in my analysis and kept only the columns that prove to be necessary in plotting the data. I want to avoid certain tokens being plotted multiple times if they are part of the same word in the same utterance.
```{r}
anger_words <- merged_data3 %>% 
  filter(Categoría=="Enojo") %>% 
  group_by(Stems) %>% 
  mutate(Freq = n()) %>% 
  ungroup() %>% 
  select(Stems, Freq, Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)

happiness_words <- merged_data3 %>% 
  filter(Categoría=="Alegría") %>% 
  group_by(Stems) %>% 
  mutate(Freq = n()) %>% 
  ungroup() %>% 
  select(Stems, Freq, Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)

sadness_words <- merged_data3 %>% 
  filter(Categoría=="Tristeza") %>% 
  group_by(Stems) %>% 
  mutate(Freq = n()) %>% 
  ungroup() %>% 
  select(Stems, Freq, Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)

surprise_words <- merged_data3 %>% 
  filter(Categoría=="Sorpresa") %>% 
  group_by(Stems) %>% 
  mutate(Freq = n()) %>% 
  ungroup() %>% 
  select(Stems, Freq, Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)

disgust_words <- merged_data3 %>% 
  filter(Categoría=="Repulsión") %>% 
  group_by(Stems) %>% 
  mutate(Freq = n()) %>% 
  ungroup() %>% 
  select(Stems, Freq, Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)

fear_words <- merged_data3 %>% 
  filter(Categoría=="Miedo") %>% 
  group_by(Stems) %>% 
  mutate(Freq = n()) %>% 
  ungroup() %>% 
  select(Stems, Freq, Categoría) %>% 
  distinct(Stems, .keep_all = TRUE)
```


## Step 8: Plotting Parliament sentence data

Generating wordclouds for each sentiment:
```{r}
#anger
set.seed(10)
wordcloud(words = anger_words$Stems, freq = anger_words$Freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.5, scale=c(3.5, 0.3),
          colors=brewer.pal(8, "Set1"))

#happiness
set.seed(10)
wordcloud(words = happiness_words$Stems, freq = happiness_words$Freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.5, scale=c(2.5, 0.3),
          colors=brewer.pal(8, "Set1"))

#sadness
set.seed(10)
wordcloud(words = sadness_words$Stems, freq = sadness_words$Freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.5, scale=c(3.5, 0.3),
          colors=brewer.pal(8, "Set1"))

#surprise
set.seed(10)
wordcloud(words = surprise_words$Stems, freq = surprise_words$Freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.5, scale=c(2, 0.3),
          colors=brewer.pal(8, "Set1"))

#disgust
set.seed(10)
wordcloud(words = disgust_words$Stems, freq = disgust_words$Freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.5, scale=c(3.5, 0.3),
          colors=brewer.pal(8, "Set1"))

#fear
set.seed(8)
wordcloud(words = fear_words$Stems, freq = fear_words$Freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.5, scale=c(3.5, 0.3),
          colors=brewer.pal(8, "Set1"))
```

Generating bar charts for each sentiment:
```{r}
library(ggplot2)

anger_words2 <- anger_words %>% 
  filter(Freq > 1)
anger_plot <- ggplot(anger_words2, aes(Stems, Freq)) + 
  geom_bar(stat = 'identity', fill="red3", width = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = 'Anger Word Frequency (n>1)')
anger_plot

happiness_words2 <- happiness_words %>% 
  filter(Freq > 10)
happiness_plot2 <- ggplot(happiness_words2, aes(Stems, Freq)) + 
  geom_bar(stat = 'identity', fill="gold1", width = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = 'Happiness Word Frequency (n>10)')
happiness_plot2
  
sadness_words2 <- sadness_words %>% 
  filter(Freq > 5)
sadness_plot <- ggplot(sadness_words2, aes(Stems, Freq)) + 
  geom_bar(stat = 'identity', fill="blue4", width = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = 'Sadness Word Frequency (n>5)')
sadness_plot

surprise_plot <- ggplot(surprise_words, aes(Stems, Freq)) + 
  geom_bar(stat = 'identity', fill="mediumspringgreen", width = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = 'Surprise Word Frequency')
surprise_plot

disgust_plot <- ggplot(disgust_words, aes(Stems, Freq)) + 
  geom_bar(stat = 'identity', fill="olivedrab1", width = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = 'Disgust Word Frequency')
disgust_plot

fear_plot <- ggplot(fear_words, aes(Stems, Freq)) + 
  geom_bar(stat = 'identity', fill="darkorange", width = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = 'Fear Word Frequency')
fear_plot
```

Finding general sentiment for each speaker:
```{r}
by_speaker <- merged_data3 %>% 
  count(speaker_id, Categoría, sort=TRUE) %>%
  filter(n == max(n), .by=speaker_id)

#Number of distinct speakers
by_speaker %>% distinct(speaker_id) %>% nrow()
```

Plotting composition of general sentiments across speakers:
```{r}
#Out of 128 general sentiments across all speakers, find sentiment distribution
#Each square represents each speaker (111 speakers total, some speakers have more than one square)
#Picture plot as a parliament session, each member represented by one square
library(waffle)

comp_waffle <- by_speaker %>%
  group_by(Categoría) %>% 
  mutate(Freq=n()) %>%
  select(Categoría, Freq) %>% 
  distinct()
comp_waffle

waffle(comp_waffle)
```
## Step 9: Comparing Parliament sentence sentiment distribution with SEL distribution

Creating waffle plot for general SEL sentiment distribution (1), creating dataframe for proportions of SEL words to sentiments (2), and creating dataframe for proportions of speakers to to sentiments
```{r}
library(knitr)
library(kableExtra)

ggplot(data = sel_distribution, aes(fill=sel_distribution$Categoría, values=Freq)) +
  geom_waffle(color = "white") +
  scale_fill_manual(values = c(Alegría = "seagreen4", Enojo = "slateblue1", 
                               Miedo = "hotpink", Repulsión = "olivedrab3", 
                               Sorpresa = "gold", Tristeza = "tomato"))

prop_words <- matrix(c((sel_distribution$Freq[1])/sum(sel_distribution$Freq),
                       (sel_distribution$Freq[2])/sum(sel_distribution$Freq),
                       (sel_distribution$Freq[3])/sum(sel_distribution$Freq),
                       (sel_distribution$Freq[4])/sum(sel_distribution$Freq),
                       (sel_distribution$Freq[5])/sum(sel_distribution$Freq),
                       (sel_distribution$Freq[6])/sum(sel_distribution$Freq)), ncol=1)
rownames(prop_words) <- c('Happiness', 'Anger', 'Fear', 'Disgust', 'Surprise', 'Sadness')
colnames(prop_words) <- c('Proportion')
kable(prop_words) %>% save_kable("prop_words.txt")

prop_usage <- matrix(c((comp_waffle$Freq[1])/sum(comp_waffle$Freq),
                       (comp_waffle$Freq[3])/sum(comp_waffle$Freq),
                       (comp_waffle$Freq[4])/sum(comp_waffle$Freq),
                       (comp_waffle$Freq[6])/sum(comp_waffle$Freq),
                       (comp_waffle$Freq[5])/sum(comp_waffle$Freq),
                       (comp_waffle$Freq[2])/sum(comp_waffle$Freq)), ncol=1)
rownames(prop_usage) <- c('Happiness', 'Anger', 'Fear', 'Disgust', 'Surprise', 'Sadness')
colnames(prop_usage) <- c('Proportion')
kable(prop_usage) %>% save_kable("usage_words.txt")

#Comparing SEL Data matrix with Speech Data matrix
prop_words - prop_usage

#Comparing Economy Reference Set Data matrix with Speech Data matrix
```

## Step 9: Comparing Parliament sentence sentiment distribution with non-related 'economy' distribution

Here, I will follow many of the steps above to create a new distribution that is based off of a brand-new keyword vector. Since all data is stemmed, the vector I chose only contains one item: 'econom'
```{r}
#Creating new dataframe just with economy-related stem
keywords_econ <- ('econom')

keywords_col_econ <- small_data_span$sentence %>% 
  str_detect(keywords_econ)

small_data_econ <- small_data_span %>%
  mutate(keywords = keywords_col_econ)

small_data_econ <- small_data_econ %>%
  filter(keywords == TRUE)

#Tokenizing sentences
small_data_econ <- small_data_econ %>%
  unnest_tokens(tokens, sentence, token="words", drop=FALSE, to_lower=TRUE)

#Stemming sentence tokens
econ_stems <- wordStem(small_data_econ$tokens, language = "spanish")
stemmed_econ_ngrams <- small_data_econ %>% 
  mutate(Stems = econ_stems)

#Inner-joining stemmed SEL set and newly stemmed economy set
merged_econ_data <- stemmed_econ_ngrams %>% 
  inner_join(stemmed_sel, by="Stems", multiple = "all", unmatched = "drop", relationship = "many-to-many")

#Finding distribution of sentiments across individual speakers
by_speaker_econ <- merged_econ_data %>% 
  count(speaker_id, Categoría, sort=TRUE) %>%
  filter(n == max(n), .by=speaker_id)

#Plotting with waffle()
econ_waffle <- by_speaker_econ %>%
  group_by(Categoría) %>% 
  mutate(Freq=n()) %>%
  select(Categoría, Freq) %>% 
  distinct()
waffle(econ_waffle)

#Gathering proportions
econ_usage <- matrix(c((econ_waffle$Freq[1])/sum(econ_waffle$Freq),
                       (econ_waffle$Freq[2])/sum(econ_waffle$Freq),
                       (econ_waffle$Freq[3])/sum(econ_waffle$Freq),
                       (econ_waffle$Freq[4])/sum(econ_waffle$Freq),
                       (econ_waffle$Freq[5])/sum(econ_waffle$Freq)), ncol=1)

rownames(econ_usage) <- c('Happiness', 'Sadness', 'Anger', 'Disgust', 'Surprise')
colnames(econ_usage) <- c('Proportion')
kable(econ_usage) %>% save_kable("econ_words.txt")
```


