---
title: "Progress-Report-1"
author: "Claire McLean"
date: "2024-10-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

NOTE: The dataset I am using for this project is named basque_parliament_1 and was found on https://huggingface.co. The original project was funded by the Spanish Ministry of Science and Innovation, and is licensed as cc0-1.0.

Citation:
@misc {software_technologies_working_group_2024,
	author       = { {Software Technologies Working Group} },
	title        = { basque_parliament_1 (Revision a2fbcaf) },
	year         = 2024,
	url          = { https://huggingface.co/datasets/gttsehu/basque_parliament_1 },
	doi          = { 10.57967/hf/2485 },
	publisher    = { Hugging Face }
}

# Progress Report 1

## Here, I will begin the next steps of my data cleaning.

## Step 1: Performed in Python
In this step, I wrote a script to filter the train.tsv file provided in the Hugging Face dataset folder to include only the rows/utterances that contain the following keywords:

['hablante', 'euskera', 'lenguaje', 'idioma', 'aprendizaje', 'lingüístico', 'lingüística']

These keywords were chosen because they demonstrate the relevance of the utterance to the concept of language, linguistics, or "speaker".

The final output of the script was a smaller .csv file containing only the rows that will be relevant to my analysis. This dataframe will be further cleaned in later steps.

This script has been included in my Progress Report 1, but will most likely not be part of my final submission.

## Step 2: Data Wrangling in R

First, import dependencies:
```{r}
library("tidyverse")
library("readr")
```


Next, read in the .csv file:
```{r}
rel_data <- read_csv("BS_RelData.csv")
```
Reducing size by only including first 200 rows:
```{r}
small_data <- rel_data %>% slice(1:250)
view(small_data)
```

Now, filtering to only include rows that are Spanish utterances (exclude Basque and bilingual):
```{r}
small_data_span <- small_data %>% filter(language == "es")
view(small_data_span)
```

Now removing the 'path' column as it is unnecessary for my analysis:
```{r}
small_data_span2 <- small_data_span %>% select(-path)
view(small_data_span2)
```

This leaves me with 190 tidy rows to annotate for sentiment. This number may change depending on the analysis process.

Now exporting dataframe to .csv file (annotations still needed):
```{r}
write_csv(small_data_span2, "small_data_NA.csv")
```

## Some basic stats:
```{r}
#number of rows:
nrow(small_data_span2)

#number of columns:
ncol(small_data_span2)

#average utterance length
mean(small_data_span2$length)

#number of unique speakers
n_distinct(small_data_span2$speaker_id)
```

